### 2019-09-04 ic-journal/part2.tex
cancelled renaming of quantifiers from E -> P/Q as now they're talking about equations too.
Added multivariate.[h]tex
Improved TeX output of "nil" (final changes on HOL side)

\paragraph{Strong bisimularity}
Based on results about ``bisimulation up to $\sim$'', we have first proved the following
 lemma. It states  that if $X$ is weakly guarded
in $E$, then the ``first move'' of $E$ is independent of the agent
substituted for $X$:
\begin{lemma}[Lemma 3.13 of \cite{Mil89}]
\label{lem:313}
If the variable $X$ are weakly guarded in $E$, and
$E\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$E'\{P/X\}$ (for some expression $E'$), and moreover, for any $Q$,
$E\{Q/X\}\overset{\alpha}{\rightarrow} E'\{Q/X\}$:
\begin{alltt}
\hfill{[STRONG_UNIQUE_SOLUTION_LEMMA]}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{E} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}. \HOLConst{CONTEXT} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{E} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

\paragraph{Weak bisimilarity}
Milner's book \cite{Mil89} contains only two ``unique solutions of
equations'' theorems, one for strong equivalence, the other for
observational congruence.   There is however a
version  of the result for weak bisimilarity that shares a large portion of proof
steps with that case for observational congruence.  
As  weak bisimilarity 
 is not congruence, we have to be more restrictive on the syntax for
 the equation, using strong guardedness.
% to make correct statement of this
% theorem, we must slightly modify the concept of sequential expresses
% with further restrictions: no direct sums (or only guarded sums).
% The ``unique solution of equations'' theorem for weak equivalence
% requires more restrictions on the expression, namely, .
% Now we're ready to state and prove the following lemma:
\begin{lemma}
If the variable $X$ is strongly guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some expression $H$, and for any $Q$, we have
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}.
           \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{a} \HOLSymConst{=} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}) \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}\hfill{[WEAK_UNIQUE_SOLUTION_LEMMA]}
\end{alltt}
\end{lemma}

We can then derive the target theorem:
\begin{theorem}{(Unique solution of equations for weak bisimilarity)}
Let $E$ be guarded and sequential expressions, and let $P \wb
E\{P/X\}$,
$Q \wb E\{Q/X\}$. Then $P \wb Q$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLBoundVar{Q}\hfill{[WEAK_UNIQUE_SOLUTION]}
\end{alltt}
\end{theorem}

@@

\begin{lemma}
% To prove two processes $E$ and $E'$  observationally congruent, it is
% sufficient enough to construct a bisimulation with additional observational
% transition properties:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WEAK_BISIM} \HOLFreeVar{Wbsm} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}.
       (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u}.
            (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenImp{} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
            \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenImp{} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
       \HOLBoundVar{E} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{E\sp{\prime}}\hfill{[OBS_CONGR_BY_WEAK_BISIM]}
\end{alltt}
\end{lemma}

This lemma plays a key role in the
corresponding `unique solutions' theorem, in connection with the
following result. 
In this part, we have not 
used 
 ``bisimulation up-to'' techniques. Using
 the  above \texttt{OBS_CONGR_BY_WEAK_BISIM},
 we  directly construct the appropriate bisimulation relations.
\begin{lemma}
If the variable $X$ is strongly guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some  $H$, and for any $Q$, we also have 
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{SEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{a} \HOLSymConst{=} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}) \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}\hfill{[OBS_UNIQUE_SOLUTION_LEMMA]}
\end{alltt}
\end{lemma}

\begin{theorem}{(Unique solution of equations for observational congruence)}
Let $E$ be guarded and sequential expressions, and let $P \rapprox
E\{P/X\}$,
$Q \rapprox E\{Q/X\}$. Then $P \rapprox Q$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{Q}\hfill{[OBS_UNIQUE_SOLUTION]}
\end{alltt}
\end{theorem}

@@

A trace is represented by the initial and final processes, plus
a list of actions  so performed.
For this, we first 
 define the concept of label-accumulated reflexive transitive closure
 (\texttt{LRTC}).
Given a labeled transition relation \texttt{R} on CCS, \texttt{LRTC R} is
a label-accumulated relation representing the trace of transitions:
\begin{alltt}
\HOLConst{LRTC} \HOLFreeVar{R} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b} \HOLSymConst{\HOLTokenDefEquality{}}
  \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P}.
      (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x}. \HOLBoundVar{P} \HOLBoundVar{x} [] \HOLBoundVar{x}) \HOLSymConst{\HOLTokenConj{}}
      (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z}. \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLBoundVar{x} (\HOLBoundVar{h}\HOLSymConst{::}\HOLBoundVar{t}) \HOLBoundVar{z}) \HOLSymConst{\HOLTokenImp{}}
      \HOLBoundVar{P} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b}\hfill{[LRTC_DEF]}
\end{alltt}
The trace relation for CCS can be then obtained
 by calling \texttt{LRTC} on the (strong, or single-step) labeled transition
 relation \texttt{TRANS} ($\overset{\mu}{\rightarrow}$) defined by SOS rules:
\begin{alltt}
\HOLConst{TRACE} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{LRTC} \HOLConst{TRANS}\hfill{[TRACE_def]}
\end{alltt}

If the list of actions is empty, that means that there is no transition and hence,
if there is at most one visible action (i.e., a label) in the list of actions,
then the trace is also a weak transition. Here
we have to distinguish between two cases: no label and unique label (in
the list of actions). The definition of ``no
label'' in an action list is easy (here \texttt{MEM} tests if a given element is a member of a list):
\begin{alltt}
\HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} (\HOLConst{label} \HOLBoundVar{l}) \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}

The definition of ``unique label'' can be done in many ways, the
following definition (a suggestion from Robert Beers)
avoids any counting or filtering in the list.
It says that a label is unique in a list of actions if and only if there is no
label in the rest of list:
\begin{alltt}
\HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}}
  \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{L\sb{\mathrm{1}}} \HOLBoundVar{L\sb{\mathrm{2}}}. \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} [\HOLFreeVar{u}] \HOLSymConst{\HOLTokenDoublePlus} \HOLBoundVar{L\sb{\mathrm{2}}} \HOLSymConst{=} \HOLFreeVar{L} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{2}}}\hfill{[UNIQUE_LABEL_def]}
\end{alltt}

>>>>

\paragraph{Strong bisimularity}
Based on results about ``bisimulation up to $\sim$'', we have first proved the following
 lemma. It states  that if $X$ is weakly guarded
in $E$, then the ``first move'' of $E$ is independent of the agent
substituted for $X$:
\begin{lemma}[Lemma 3.13 of \cite{Mil89}]
\label{lem:313}
If the variable $X$ are weakly guarded in $E$, and
$E\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$E'\{P/X\}$ (for some expression $E'$), and moreover, for any $Q$,
$E\{Q/X\}\overset{\alpha}{\rightarrow} E'\{Q/X\}$:
\begin{alltt}
\hfill{[STRONG_UNIQUE_SOLUTION_LEMMA]}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}. \HOLFreeVar{E} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}. \HOLConst{CONTEXT} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{E} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

\paragraph{Weak bisimilarity}
Milner's book \cite{Mil89} contains only two ``unique solutions of
equations'' theorems, one for strong equivalence, the other for
observational congruence.   There is however a
version  of the result for weak bisimilarity that shares a large portion of proof
steps with that case for observational congruence.  
As  weak bisimilarity 
 is not congruence, we have to be more restrictive on the syntax for
 the equation, using strong guardedness.
% to make correct statement of this
% theorem, we must slightly modify the concept of sequential expresses
% with further restrictions: no direct sums (or only guarded sums).
% The ``unique solution of equations'' theorem for weak equivalence
% requires more restrictions on the expression, namely, .
% Now we're ready to state and prove the following lemma:
\begin{lemma}
If the variable $X$ is strongly guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some expression $H$, and for any $Q$, we have
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}\hfill{[WEAK_UNIQUE_SOLUTION_LEMMA]}
\end{alltt}
\end{lemma}

We can then derive the target theorem:
\begin{theorem}{(Unique solution of equations for weak bisimilarity)}
Let $E$ be guarded and sequential expressions, and let $P \wb
E\{P/X\}$,
$Q \wb E\{Q/X\}$. Then $P \wb Q$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLFreeVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Q} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLFreeVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{Q}\hfill{[WEAK_UNIQUE_SOLUTION]}
\end{alltt}
\end{theorem}

@@

\begin{lemma}
% To prove two processes $E$ and $E'$  observationally congruent, it is
% sufficient enough to construct a bisimulation with additional observational
% transition properties:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WEAK_BISIM} \HOLFreeVar{Wbsm} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}.
       \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u}.
            \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
            \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLBoundVar{E} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{E\sp{\prime}}\hfill{[OBS_CONGR_BY_WEAK_BISIM]}
\end{alltt}
\end{lemma}

This lemma plays a key role in the
corresponding `unique solutions' theorem, in connection with the
following result. 
In this part, we have not 
used 
 ``bisimulation up-to'' techniques. Using
 the  above \texttt{OBS_CONGR_BY_WEAK_BISIM},
 we  directly construct the appropriate bisimulation relations.
\begin{lemma}
If the variable $X$ is strongly guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some  $H$, and for any $Q$, we also have 
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{SEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}\hfill{[OBS_UNIQUE_SOLUTION_LEMMA]}
\end{alltt}
\end{lemma}

\begin{theorem}{(Unique solution of equations for observational congruence)}
Let $E$ be guarded and sequential expressions, and let $P \rapprox
E\{P/X\}$,
$Q \rapprox E\{Q/X\}$. Then $P \rapprox Q$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLFreeVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Q} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLFreeVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{Q}\hfill{[OBS_UNIQUE_SOLUTION]}
\end{alltt}
\end{theorem}

@@

A trace is represented by the initial and final processes, plus
a list of actions  so performed.
For this, we first 
 define the concept of label-accumulated reflexive transitive closure
 (\texttt{LRTC}).
Given a labeled transition relation \texttt{R} on CCS, \texttt{LRTC R} is
a label-accumulated relation representing the trace of transitions:
\begin{alltt}
\HOLConst{LRTC} \HOLFreeVar{R} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b} \HOLSymConst{\HOLTokenDefEquality{}}
  \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P}.
      \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x}. \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{[}\ensuremath{]} \HOLBoundVar{x}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z}. \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{(}\HOLBoundVar{h}\HOLSymConst{::}\HOLBoundVar{t}\ensuremath{)} \HOLBoundVar{z}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
      \HOLBoundVar{P} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b}\hfill{[LRTC_DEF]}
\end{alltt}
The trace relation for CCS can be then obtained
 by calling \texttt{LRTC} on the (strong, or single-step) labeled transition
 relation \texttt{TRANS} ($\overset{\mu}{\rightarrow}$) defined by SOS rules:
\begin{alltt}
\HOLConst{TRACE} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{LRTC} \HOLConst{TRANS}\hfill{[TRACE_def]}
\end{alltt}

If the list of actions is empty, that means that there is no transition and hence,
if there is at most one visible action (i.e., a label) in the list of actions,
then the trace is also a weak transition. Here
we have to distinguish between two cases: no label and unique label (in
the list of actions). The definition of ``no
label'' in an action list is easy (here \texttt{MEM} tests if a given element is a member of a list):
\begin{alltt}
\HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} \ensuremath{(}\HOLConst{label} \HOLBoundVar{l}\ensuremath{)} \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}

The definition of ``unique label'' can be done in many ways, the
following definition (a suggestion from Robert Beers)
avoids any counting or filtering in the list.
It says that a label is unique in a list of actions if and only if there is no
label in the rest of list:
\begin{alltt}
\HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{L\sb{\mathrm{1}}} \HOLBoundVar{L\sb{\mathrm{2}}}. \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} \ensuremath{[}\HOLFreeVar{u}\ensuremath{]} \HOLSymConst{\HOLTokenDoublePlus} \HOLBoundVar{L\sb{\mathrm{2}}} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{L} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{2}}}\hfill{[UNIQUE_LABEL_def]}
\end{alltt}

The final relationship between traces and weak transitions is stated
and proved in the following theorem
(where the  variable $acts$ stands for
a list of actions); 
it says, a weak transition $P\overset{u}{\Rightarrow}P'$ is also a
trace $P\overset{acts}{\longrightarrow}P'$ with a
 non-empty action list $acts$, in which either there is no label (for $u = \tau$), or 
$u$ is the unique label (for $u \neq \tau$):
%\begin{lemma}
% A weak transition $P\overset{u}{\Rightarrow}P'$ is a just trace with non
% empty action list: 1) without any visible label, if $u = \tau$, or 2)
% $u$ is the unique label in the list, if $u \neq \tau$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{P} \HOLTokenWeakTransBegin\HOLFreeVar{u}\HOLTokenWeakTransEnd \HOLFreeVar{P\sp{\prime}} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{acts}.
       \HOLConst{TRACE} \HOLFreeVar{P} \HOLBoundVar{acts} \HOLFreeVar{P\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{NULL} \HOLBoundVar{acts} \HOLSymConst{\HOLTokenConj{}}
       \HOLKeyword{if} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLKeyword{then} \HOLConst{NO_LABEL} \HOLBoundVar{acts} \HOLKeyword{else} \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLBoundVar{acts}\hfill{[WEAK_TRANS_AND_TRACE]}
\end{alltt}
%\end{lemma}
